---
title: "The 'giacR' package"
output: github_document
---

*R interface to 'Giac'*

___

Giac is a general purpose symbolic algebra software. It powers the graphical 
interface Xcas. This package allows to execute Giac commands in R. You can 
find the [documentation of Giac here](https://www-fourier.ujf-grenoble.fr/~parisse/giac/doc/en/cascmd_en/cascmd_en.html).

## Installation

```r
remotes::install_github("stla/giacR")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```


## Initialisation of a Giac session

The 'chromote' package is used to create a Giac session. If the `find_chrome()` 
function of 'chromote' returns `NULL`, you can set the path to the Chrome 
executable (or Chromium, Brave, etc) to the environment variable 
`CHROMOTE_CHROME`. Or you can pass it to the the `Giac$new` function. 
Since the Chrome executable is in my system path, I can use 
`Sys.which("chrome")`.

```{r giacNew}
library(giacR)
giac <- Giac$new(Sys.which("chrome"))
```

## Examples 

### Elementary calculus

```{r}
giac$execute("2 + 3/7")
```


### Gröbner bases

```{r}
giac$execute("gbasis([x^3 - 2*x*y, x^2*y - 2*y^2 + x], [x, y])")
```


### Close session

```{r}
giac$close()
```

